---
title: "D-RUG 2017-10-26 Map/Walk Functions in purrr"
author: "Brandon Hurr"
output: html_notebook
---

# Let's load tidyverse
``` {r setup}
library(tidyverse)
```


# R Data types
```{r vector intiation}
#initiate blank vectors of a specific type
#this can be useful when capturing data from for loops
a <- vector("logical", 5)
a
b <- vector("integer", 4)
b
c <- vector("double", 2) # can also use "numeric"
c
d <- vector("character", 10)
d
e <- vector("complex", 7)
e
f <- vector("raw", 3)
f
g <- list()
g
```
# NAs
```{r NA NA NA NA NA NA NA NA Batman}
#what do the NAs look like?
list(NA, NA_integer_, NA_real_, NA_character_, NA_complex_)
#not very exciting
```

# The NULL element

``` {r NULLs}
# NULL is the absence of a vector
a <- c(TRUE, FALSE, TRUE, FALSE, FALSE, NA) # logical vector 
a
a <- NULL
a

mtcars
mtcars[1] <- NULL #this gets rid of the mpg coloumn completely
mtcars

data(mtcars) # restore mtcars

```

# Assembling vectors into a list, dataframes and tibbles
```{r Vectors Victor, error = TRUE}
# let's put actual values into these
a <- c(TRUE, FALSE, TRUE, FALSE, FALSE, NA) #seems logical 
b <- c(1:7, NA_integer_) #Integers
c <- c(rnorm(8), NA_real_) #pull doubles from normal distribution
d <- c(LETTERS[1:3], NA_character_) #first 5 capital letters

e <- list(a, b, c, d) #make a list

e # check it twix

#Lists can contain anything of any length
#data.frame is a special type of list; things need to line up

data.frame(a, b) # uh oh

a <- c(TRUE, FALSE, TRUE, FALSE, FALSE, NA, NA, NA) # make it 8 like b
df <- data.frame(a = a, b = b) # OH YEAH /Koolaidman.gif
df 
#tibbles are a strict type of data.frame
as_tibble(df)
# with slightly more informative printing
```


# Accessing elements inside vectors, lists, dataframes, and tibbles
``` {r list_df_tibble_subsetting, error = TRUE} 
e <- list(a = a, b= b, c = c, d = d, e = list(1, "A")) #make a named list

#vectors
a <- c(TRUE, FALSE, TRUE, FALSE, FALSE, NA) #logic vector 
a[1] 
a[[1]] # does the same thing
a[[[1]]] #lolwut are you doing?


# single brackets returns a list with that element in it
# can reference by location or by name
e[1]
e["a"]

# double brackets return a vector 
e[[2]]
e[["b"]]

#stack [['s it to peel the onion back
e[["e"]][[2]]

#dolla dolla bills y'all (works liks [[, with named vectors)
e$c

#same for dataframes
mtcars[1]
mtcars[[1]]

# and for tibbles
tbbl <- tibble(x = 1:3, y = LETTERS[1:3])
tbbl[1]
tbbl["x"]

tbbl[[2]]
tbbl[["y"]]
```

# The very basic basics of functions

``` {r functionbasics}
#basic struction is assignment, definition, arguments, body, output
addxy <- function(x, y) {
  out <- x+y
  out
}

formals(addxy)
body(addxy)
environment(addxy)

addxy(1,2)

# anonymous functions are not assigned, but are used in place
# we will see this more later with dplyr and purrr
function(x) {x+1}

```

``` {r Y_make_a_function}
#Simple dataframe
df <- data.frame(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
# What are we doing here?
df$a <- (df$a - min(df$a, na.rm = TRUE)) /  
  (max(df$a, na.rm = TRUE) - min(df$a, na.rm = TRUE))
df$b <- (df$b - min(df$b, na.rm = TRUE)) /  
  (max(df$b, na.rm = TRUE) - min(df$b, na.rm = TRUE))
df$c <- (df$c - min(df$a, na.rm = TRUE)) /  
  (max(df$c, na.rm = TRUE) - min(df$c, na.rm = TRUE))
df$d <- (df$d - min(df$d, na.rm = TRUE)) /  
  (max(df$d, na.rm = TRUE) - min(df$d, na.rm = TRUE))

```


``` {r function01, error = TRUE, eval = FALSE}
df$d <- (df$d - min(df$d, na.rm = TRUE)) /  
  (max(df$d, na.rm = TRUE) - min(df$d, na.rm = TRUE))
```


``` {r function02, error = TRUE, eval = FALSE}
# pull out what's common
x <- (x - min(x, na.rm = TRUE)) /  
  (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))
```


``` {r function03, error = TRUE, eval = FALSE}

#simplify more and refactor if that makes sense
x <- (x - min(x, na.rm = TRUE)) /  
  diff(range(x, na.rm = TRUE))
```

--
``` {r function04, error = TRUE, eval = FALSE}
# rewrite as a function
rescale_0_1 <- function(x) {
  (x - min(x, na.rm = TRUE)) /  
  diff(range(x, na.rm = TRUE))
}
```

``` {r test function}
rescale_0_1 <- function(x) {
  (x - min(x, na.rm = TRUE)) /  
  diff(range(x, na.rm = TRUE))
}

testvec <- 1:11
testvec
rescale_0_1(testvec)
```

``` {r use_function}
df <- data.frame(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
rescale_0_1 <- function(x) {
  (x - min(x, na.rm = TRUE)) /  
  diff(range(x, na.rm = TRUE))
}

out <- df
out$a <- rescale_0_1(df$a)
out$b <- rescale_0_1(df$b)
out$c <- rescale_0_1(df$c)
out$d <- rescale_0_1(df$d)
out
```

# Reduce iteration with a for loop
``` {r forloops}
out <- df #make a copy to store scaled data
for (i in 1:ncol(df)) { # loop through each element
  out[i] <- rescale_0_1(df[[i]]) #apply function and store it to out
}
out
```

# seq_along

``` {r for_seq_along_loops}
out <- df #make a copy to store scaled data
for (i in seq_along(df)) {
  out[i] <- rescale_0_1(df[[i]])
}
out
```

# map to iterate
```{r WHAT_IS_THIS_MAPGIC}
map(df, function(x) rescale_0_1(x)) %>% #map over the columns in df
```
# let's look at another more simple example
# This list contains two integer vectors and we want the mean of each one
``` {r simplemap}
l = list(a=1:10, b = 10:100)
map(l, function(x) mean(x, na.rm = TRUE))
```

# We know we want one double back from each list
# specify and you get a vector back instead of a list
``` {r rightmap}
l = list(a=1:10, b = 10:100)
map_dbl(l, function(x) mean(x, na.rm = TRUE))
```

# These are type-strict functions though
# get it wrong and you'll know
``` {r wrongmap, error = TRUE}
l = list(a=1:10, b = 10:100)
map_lgl(l, function(x) mean(x, na.rm = TRUE))
```

# Map Examples

